Global patterns of forest autotrophic carbon fluxes

### 2019-09-17 manuscript/manuscript draft.Rmd

@@

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower. Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~. In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation. This pattern was also seen in the $R^2$ values for multivariate models.

*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP. MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables. MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig. 5). We found no support for a saturation point of productivity with temperature.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was evidence that productivity begins to decrease at values above 1800mm PET.

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig. 5). ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it. Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

In contrast, there was no significant effect of precipitation seasonality on productivity.

>>>>

bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up! (Extensive citations from EC literature for example.)

@@

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower. Climate explained at most 24% of the variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~. In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation. This pattern was also seen in the $R^2$ values for multivariate models.

*How does productivity relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables. MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig. 5). We found no support for a saturation point of productivity with temperature.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that productivity begins to decrease at values above 1800mm PET.

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (fig. 5). ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it. Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on productivity.
